### 2023-06-11 Article.Rmd
12.06.2023

---
title             : "From Likes to Change: Assessing the Impact of Citizen Engagement on the European Commission's Social Media Platforms"
shorttitle        : "Impact of Citizen Engagement on the European Commission's Social Media Platforms"

author: 
  - name          : "Tanase Tasențe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "tanase.tasente@365.univ-ovidius.ro"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Equal contribution"
      - "Conceptualization"
      - "Data curation"
      - "Investigation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Cristian Opariuc-Dan"
    affiliation   : "1,2"
    role:
      - "Equal contribution"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Data curation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
  - name          : "Cristina Dana Popescu"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Professional proofreading"

affiliation:
  - id            : "1"
    institution   : "Ovidius University of Constanța, Romania"
  - id            : "2"
    institution   : "University of Bucharest, Romania"

authornote: |
  Department of Administrative and Social Sciences

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib", "references.bib"]
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
documentclass     : "apa7"
classoption       : "man"
csl               : apa.csl
output            :
  pdf_document    : papaja::apa6_pdf
  word_document   : papaja::apa6_docx
header-includes:
  - \usepackage{ragged2e}
---

@@

The novelty of our study lies in the fact that (1) **[TASE]**

@@

We further performed a sentiment analysis both on data with (see Fig. \@ref(fig:boxplot-EXT) and Fig. \@ref(fig:sent-EXT)) and without extreme outliers (see Fig. \@ref(fig:boxplot-NEXT) and Fig. \@ref(fig:sent-NEXT)) and the information suggested that ... **[TASE]**

@@

# Discussion

>>>>

---
title             : "From Likes to Change: Assessing the Impact of Citizen Engagement on the European Commission's Social Media Platforms"
shorttitle        : "Impact of Citizen Engagement on the European Commission's Social Media Platforms"
author: 
  - name          : "Tanase Tasențe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "tanase.tasente@365.univ-ovidius.ro"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Equal contribution"
      - "Conceptualization"
      - "Data curation"
      - "Investigation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Cristian Opariuc-Dan"
    affiliation   : "1,2"
    role:
      - "Equal contribution"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Data curation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
  - name          : "Cristina Dana Popescu"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Professional proofreading"
affiliation:
  - id            : "1"
    institution   : "Ovidius University of Constanța, Romania"
  - id            : "2"
    institution   : "University of Bucharest, Romania"
authornote: |
  Department of Administrative and Social Sciences
abstract: |
  This study investigates the role of emotional resonance in social media communications from the European Commission (EC) and its impact on public engagement across different platforms (Facebook, Instagram, Twitter, and YouTube). Our results indicate that messages with emotional resonance significantly boost public engagement levels, with the effects varying based on the specific platform. We further reveal a notable influence of both positive and negative emotional resonance on engagement, with the latter exerting a stronger impact. This research expands our understanding of engagement dynamics on different platforms, stressing the importance of tailored platform-specific strategies. The findings underscore the crucial role of social media in enhancing public involvement in policy dialogue and emphasize the need for public institutions to incorporate emotional resonance into their communication strategies, accounting for unique platform characteristics. These insights provide a foundation for future discussions and research into optimizing communication strategies within a multi-platform environment, which will shape the future of digital governance and public policymaking.
keywords          : "social media, public institutions, emotional resonance, public engagement, platform-specific strategies, digital governance"
wordcount         : "X"
bibliography      : ["r-references.bib", "references.bib"]
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
documentclass     : "apa7"
classoption       : "man"
csl               : apa.csl
output            :
  pdf_document    : papaja::apa6_pdf
  word_document   : papaja::apa6_docx
header-includes:
  - \usepackage{ragged2e}
---

@@

The novelty of our study lies in the fact that we have systematically compared and contrasted user engagement and emotional resonance across multiple social media platforms utilized by the European Commission. While previous research has generally focused on the impact of citizen engagement within a single social media platform, our study is unique in its comprehensive cross-platform analysis.

In this context, we have pursued an integrated approach to understanding how emotional resonance can influence public engagement levels across these platforms. Furthermore, we have explored the moderating effect of the social media platform itself on the impact of emotional resonance, a novel aspect that has not been fully addressed in previous research.

The innovative approach of this study allowed us to draw robust conclusions about the nuanced relationship between communication strategy, emotional resonance, platform choice, and public engagement. Our research has therefore not only provided key insights into the patterns of citizen engagement on European Commission's social media platforms, but also presented an advanced framework for understanding how such engagement can be effectively harnessed and maximized for public outreach and policymaking.

Ultimately, this study has expanded the current discourse in the field of digital communication and citizen engagement, and opened up new avenues for future research on optimizing communication strategies in a multi-platform social media environment.

@@

A comprehensive sentiment analysis was conducted on data, both with (see Fig. \@ref(fig:boxplot-EXT) and Fig. \@ref(fig:sent-EXT)) and without extreme outliers (see Fig. \@ref(fig:boxplot-NEXT) and Fig. \@ref(fig:sent-NEXT)) suggesting distinctive tonal variations in the discourse employed by the European Commission across different social media platforms.
In the case of Facebook, the sentiment indicator's mean value was determined to be 0.14, indicative of a marginally positive tone. The sentiment ranged between -0.27 and 0.63. The discourse with a negative connotation averaged at -0.08, while the discourse with a positive undertone demonstrated a mean of 0.22, thereby further substantiating a propensity for positivity.

Akin to Facebook, the sentiment indicator on Instagram averaged at 0.12, which points towards a mildly positive sentiment, with the spectrum extending from -0.22 to 0.46. The mean sentiment score for negatively perceived discourse was -0.07, and the positively perceived discourse showed an average of 0.20, once again suggesting a bend towards positive discourse on this platform.

Conversely, Twitter displayed a broader sentiment spectrum. Although the sentiment indicator mean was marginally higher than Facebook and Instagram at 0.15, it ranged from -0.63 to 1.00, implying a more diverse expression of sentiments. Negative discourse manifested an average of -0.07, identical to Instagram but more negative than Facebook. Interestingly, Twitter maintained the highest average score for positive discourse, at 0.22, equal to the positivity in Facebook and surpassing that on Instagram.

On YouTube, the sentiment indicator's mean value was observed to be the lowest amongst the evaluated platforms at 0.10, spanning from -0.19 to 0.34, thereby indicating a more tempered sentiment. Negative discourse on YouTube scored the lowest mean of -0.03, while positive discourse registered the minimum average of 0.13 compared to the other platforms.

In conclusion, despite a general inclination towards a slightly positive mean sentiment across all platforms, Twitter demonstrated the most substantial sentiment range, signifying the potential for both intensely negative and positive discourse. Instagram and Facebook depicted a modestly positive sentiment with less variability, while YouTube exhibited the most tempered sentiment range, with the lowest averages for both positive and negative discourse.

@@

# Discussion
This investigation commenced with the aim of assessing the impact of citizen engagement on the European Commission's social media platforms by specifically evaluating the influence of emotionally resonant communications on public engagement levels. The discussion here explores the findings derived from the study and their implications, juxtaposing them against the extant body of research on social media engagement, public institutional communication, and the role of emotional resonance in these facets.

The first hypothesis (H1) was predicated on the notion that communications with emotional resonance would yield higher public engagement levels. The research findings provided empirical support to this hypothesis, demonstrating a significant association between emotionally resonant communications and amplified levels of public engagement. This echoes previous scholarly work emphasizing the role of emotional appeal in enhancing audience engagement on social media platforms (Dolan et al., 2016; Smith & Gallicano, 2015). However, our research contributes a new dimension by bringing these aspects into the realm of public institutions, specifically the European Commission.

Public institutions often struggle with perceived remoteness and disconnection from citizens. However, the advent of social media offers a new avenue for these institutions to reach out to their citizenry, thereby allowing them to foster greater trust and engagement. Our findings confirm that the integration of emotional resonance within these online communications could significantly enhance this engagement process, fostering a more meaningful and interactive dialogue between the institutions and citizens.

Our findings align with the theoretical underpinnings of the uses and gratifications theory, emphasizing that emotional resonance serves as a primary motivation for users to engage with social media content. Consequently, public institutions should incorporate emotionally resonant messages into their social media communications to meet these user motivations, thus promoting higher engagement levels. This reflects the findings of Dolan et al. (2016), who identified emotional resonance as a key motivator in user engagement behavior.

The second hypothesis (H2) postulated that the social media platform could exert an influence on the emotional resonance in public engagement. The findings of the study substantiated this claim, revealing that the type of social media platform indeed moderates the impact of emotional resonance on public engagement. This variability across platforms underscores the need for a more nuanced, platform-specific approach to communication strategies.

Such an approach resonates with the findings of @Voorveld2018 who noted the differential roles of different social media platforms in user engagement. Public institutions, therefore, need to be mindful of these variations, ensuring that their communication strategies are tailored to the specific features, functionalities, and user behaviors of each platform to maximize the effectiveness of their emotionally resonant communications.

Our third hypothesis (H3) proposed that the social media platform could moderate the effect of emotional resonance on public engagement. This hypothesis was also validated by the study, underscoring the pivotal role that the choice of the social media platform plays in shaping public engagement levels. Each platform presents unique attributes and user behaviors, influencing the extent to which emotional resonance can stimulate public engagement.

This finding also converges with the research conducted by @Voorveld2018, which asserted the differentiating impact of platform type on social media engagement. Hence, the insights derived from our study reinforce the need for public institutions to strategically consider their choice of social media platform, factoring in the distinctive characteristics of each platform and their corresponding audience base to optimize their emotionally resonant communications.

A particularly interesting revelation was the relative impact of positive and negative emotional resonance on engagement levels. Both types of emotional resonance were found to amplify engagement levels, yet negative emotional resonance appeared to exert a stronger influence. This suggests that audiences may be more inclined to engage with messages that evoke negative emotions. Public institutions must carefully navigate this complex terrain, ensuring that their communications strike a delicate balance between positive and negative emotional resonance, taking into consideration the potential risks and rewards of each approach.

In sum, this study provides a wealth of empirical evidence to support the significance of emotional resonance in amplifying citizen engagement on the social media platforms of public institutions. It is a clarion call for public organizations to strategically incorporate emotional resonance into their communications, thereby leveraging this engagement driver to foster a more interactive, engaging dialogue with their audiences. Further, it highlights the pivotal role of platform choice, emphasizing the need to take into account the unique attributes and audiences of each platform in formulating and executing effective communication strategies.

However, despite its contributions, this study is not without its limitations. Future research should delve deeper into the intricate relationship between emotional resonance and public engagement, examining potential moderating factors such as the type of communication, audience demographics, or the cultural context. Additionally, further studies could evaluate the differential impact of positive and negative emotional resonance, providing a more nuanced understanding of how these contrasting types of emotional appeal influence public engagement. The role of individual social media platforms and their unique features in shaping emotional resonance could also be explored in more detail, providing insights into how public institutions can optimize their platform-specific strategies.

This research offers a critical springboard for public institutions to understand and leverage emotional resonance in their social media communications, thereby unlocking new opportunities for more effective and engaging citizen interactions. As public institutions continue to navigate the digital landscape, they must take these findings into account, recognizing the crucial role of emotional resonance in social media communications as well as the importance of adopting a platform-specific approach. The benefits of doing so will not only enhance public engagement but also foster a stronger, more trustful relationship between public institutions and their audiences.
